---
title: "Assignment 1"
author: "Ruijie Sun"
date: "January 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### 1. (30 marks) Consider the box plot below, drawn in R, based on the data in the ???le "juries.csv".

#### (a) (10 marks) Recall the 1.5IQR Rule which is used to identify potential outliers. Show, using this rule, how the two points identify as outliers. 

*Appendix:*

```{r}
# Import the data
dat_juries <- read_csv("juries(1).csv")

# preview the top 5 rows
dat_juries %>% head(5)


```

```{r}
# compute the IQR
as.numeric(dat_juries %>% filter(JUDGE=="SPOCKS") %>% select(PERCENT)

```



*Solution:*




#### (b) (15 marks) Recreate the side-by-side box plots of percent of women on venires for Spock's judge and the other judges without identifying outliers. 

#### (c) (5 marks) Comment on the difference between the skeletal box plot (which does not identify outliers) and the modi???ed box plot (which identi???es outliers).
