\documentclass[11pt]{article}
\usepackage{fullpage}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{clrscode3e}

\setlength{\parskip}{2ex}
\newcommand{\qedsymb}{\hfill{\rule{2mm}{2mm}}}
\newenvironment{proof}{\vspace{-1mm}\begin{trivlist}
}{\qedsymb\end{trivlist}\vspace{-1mm}}

\begin{document}
\begin{center}
{\bf \Large \bf CSC373 Winter 2015 Problem Set \# 3}\\
Name: Weidong An\\
Student Number: 1000385095\\
UTOR email: weidong.an@mail.utoronto.ca\\
\today\\
\end{center}

\begin{enumerate}[label=(\alph*)]
\item
Subproblems: Determine whether or not you can purchase exactly $N-3$ or $N-10$ or $N-25$ Tofu Nuggets given that
you can only purchase boxes of 3, 10 or 25.\\
Solution to the problem: If one of three subproblems is true, then the solution to this problem is true. Otherwise, it is false.

Suppose you can purchase exactly $N$ Tofu Nuggets. Then "last" box of Tofu Nuggets purchased is of size 3 or 10 or 25. So the problem can be reduced to whether or not you can purchase exactly $N-3$ or $N-10$ or $N-25$ Tofu Nuggets.

\item
Define array $T$ such that:\\
The indices of $T$ are from $-24$ to $N$.\\
Each element in $T$ is the boolean value $\const{true}$ or $\const{false}$\\
$T[0] = \const{true}$\\
$T[i] = \const{false}$ for $-24\leq i \leq -1$ \\
$T[i] = T[i-3]$ or $T[i-10]$ or $T[i-25]$ for $i > 0$\\

\item
Base cases: $T[0]=\const{true}$ and $T[i]=\const{false}$ for $ -24\leq i \leq -1$.\\
General cases: $T[i] = T[i-3]$ or $T[i-10]$ or $T[i-25]$ for $i >0$

\item

\begin{codebox}
\li \For $i$ in $-24, ..., -1$ 
\li      \Then $T[i] \gets \const{false}$\End
\li $T[0] \gets \const{true}$
\li \For $i$ in $1,..., N$
\li      \Then $T[i]\gets T[i-3]$ or $T[i-10]$ or $T[i-25]$\End
\li \Return $T[N]$
\end{codebox}

Worst case complexity: $\Theta(N)$\\
It is in polynomial time.\\
Explanation: The first loop (line 1 to line 2) takes 24 steps which is constant time. Line 3 takes constant time to run. The second loop (line 4 to line 5) takes $(3+1)N = 4N$ (3 for accessing array indices and 1 for getting the boolean value of "or"). Therefore in total, the algorithm takes $\Theta(N)$ time to run. It is in polynomial time.

\end{enumerate}

\end{document}